---
title: "Lab 10"
author: "Stephanie Lee"
date: "`r Sys.Date()`"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSQLite)
library(DBI)

# Initialize a temp in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)

dbListTables(con)
```
```{sql, connection=con}
PRAGMA table_info(actor)
```
# Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.
```{sql connection=con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
```

# Exercise 2
Retrieve the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{sql connection=con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
```

# Exercise 3

Write a query against the `rental` table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.

```{sql connection=con}
PRAGMA table_info(rental)
```
```{sql connection=con}
SELECT DISTINCT customer_id, rental_date
FROM rental
WHERE date(rental.rental_date) = '2005-07-05'
```

# Exercise 4
```{sql connection=con}
PRAGMA table_info(payment)
```

## Exercise 4.1

Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.

```{sql connection=con}
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
```

## Exercise 4.2

Construct a query that retrives all rows from the payment table where the amount is greater then 5

```{sql connection=con}
SELECT *
FROM payment
WHERE amount > 5
```

## Exercise 4.3

Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8

```{sql connection=con}
SELECT *
FROM payment
WHERE amount > 5 AND amount < 8
```

# Exercise 5

Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.

```{sql connection=con}
PRAGMA table_info(customer)
```

```{sql connection=con}
SELECT c.customer_id, p.payment_id, p.amount
FROM customer AS c INNER JOIN payment AS P
  ON c.customer_id = p.customer_id
WHERE c.last_name IN ('DAVIS')
```

# Exercise 6

## Exercise 6.1

Use COUNT(*) to count the number of rows in rental

```{r}
dbGetQuery(con, "
SELECT COUNT(*) AS count
FROM rental
")
```
```{sql connection=con}
SELECT COUNT(*) AS count
FROM rental
```

## Exercise 6.2

Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id

```{sql connection=con}
SELECT customer_id, COUNT(*) AS count
FROM rental
GROUP BY customer_id
```

## Exercise 6.3

Repeat the previous query and sort by the count in descending order

```{sql connection=con}
SELECT customer_id, COUNT(*) AS count
FROM rental
GROUP BY customer_id
ORDER BY count DESC
```

## Exercise 6.4

Repeat the previous query but use HAVING to only keep the groups with 40 or more.

```{sql connection=con}
SELECT customer_id, COUNT(*) AS count
FROM rental
GROUP BY customer_id
HAVING count >= 40
ORDER BY count DESC
```

# Exercise 7

The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM

## Exercise 7.1

Modify the above query to do those calculations for each customer_id

```{sql connection=con}
SELECT customer_id,
       MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       AVG(amount) AS avgpayment,
       SUM(amount) AS sumpayment
FROM payment
GROUP BY customer_id
```

## Exercise 7.2
```{sql connection=con}
SELECT customer_id,
       COUNT(*) N,
       MAX(amount) AS maxpayment,
       MIN(amount) AS minpayment,
       AVG(amount) AS avgpayment,
       SUM(amount) AS sumpayment
FROM payment
GROUP BY customer_id
HAVING N > 5
```


# Cleanup
Run the following chunk to disconnect from the connection.
```{r cleanup}
dbDisconnect(con)
```
