---
title: "Lab 9"
author: "Stephanie Lee"
date: "`r Sys.Date()`"
output: 
  github_document:
    html_preview: false
  html_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
Three examples of problems that may be solved using parallel computing 
1. Can run multiple tests at once - multtest package
2. Running multiple simulations at once

# Problem 2
The following functions can be written to be more efficient without using parallel:
\

### 1. This function generate a `n x k` dataset with all its entries distributed poisson with mean `lambda`

```{r fun1}
set.seed(1235)

# This function creates 100 vectors of length 4
fun1 <- function(n = 100, k = 4, lambda = 4) {
  x <- NULL
  
  for (i in 1:n)
    x <- rbind(x, rpois(k, lambda))
  
  return(x)
}

# Create one matrix instead of individual vectors
fun1alt <- function(n = 100, k = 4, lambda = 4) {
  
  x <- matrix(rpois(n*k,lambda), ncol=4)
  
  return(x)
}

# Benchmarking
microbenchmark::microbenchmark(
  fun1(),
  fun1alt()
)
```
We can see that generating the matrix `fun1alt` is way faster than generating by four numbers at a time `fun1()`

\
### Find the column max with `max.col()`

```{r maxcol}
# Data Generating Process
set.seed(1234)
M <- matrix(runif(12), ncol=4)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)
}
fun2(x=M)

fun2alt <- function(x) {
  idx <- max.col(t(x))
  x[cbind(idx, 1:4)]
}
fun2alt(x=M)

x <- matrix(rnorm(1e4), nrow=10)

# Benchmarking
microbenchmark::microbenchmark(
  fun2(x),
  fun2alt(x)
)
```
We see that using `max.col()` is way faster than individually applying a max function.
\


