---
title: "Lab 3"
author: "Stephanie Lee"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Initial Question: We want to pick out the weather station with maximum elevation and examine its windspeed and temperature.


### 1. Read in the data
```{r}
if(!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```

### 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
head(met)
tail(met)
```
### 3. Take a look at the variables
```{r}
str(met)
```

### 4. Take a closer look at the key variables
```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
```

```{r}
summary(met$elev)
```

```{r}
summary(met$wind.sp)
```

Looks like elevation variable has obs with values 9999 (not reasonable) so likely an indicator for missing values. Reassign 9999 vals as NA
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

The weather station with the highest elevation is at `r max(met$elev, na.rm=TRUE)` meters.

The minimum temp is also -40C (not reasonable), so remove those vals also.
```{r}
met <- met[temp > -40]
met2 <- met[order(temp)]
head(met2)
```
Still a -17.2C temp that seems suspicious.

### 5. Check the data against an external source.
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```
### 6. Calculate summary statistics keeping the initial question in mind.
Summaries at the highest elevation:
```{r}
elev <- met[elev==max(elev)]
summary(elev)
```

```{r}
met[elev==max(elev, na.rm=TRUE), .(
  temp_wind = cor(temp, wind.sp, use = "complete"),
  temp_hour = cor(temp, hour, use = "complete"),
  wind.sp_day = cor(wind.sp, day, use = "complete"),
  wind.sp_hour = cor(wind.sp, hour, use = "complete")
)]
```

### 7. Exploratory graphs
Look at distributions of all key variuable to make syre there are no remaining issues with the data.
```{r}
hist(met$elev, breaks=100)
```
```{r}
hist(met$temp)
```
```{r}
hist(met$wind.sp)
```
Late for analysis, consider doing a log transform on wind speed and elevation since they are very skewed.

Where is the weather station with the highest elevation located?
```{r}
library(leaflet)
library(webshot)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

Highest elevation is located in the Colorado wilderness west of Denver.

Plot the time series of temp and wind speed.
```{r}
```

