---
title: "README"
author: "Stephanie Lee"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(lubridate)
```

# 1. Read in data
```{r read data}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../Lab 3/met_all.gz")
```
# 2. Prep Data

Remove temps less than -17 and change elev 9999 to missing value.

```{r remove vals}
met <- met[temp > -17][elev == 9999.0, elev := NA]
summary(met)
```
Generate a date variable:

```{r date}
met <- met[ , ymd := as.Date(paste(year, month, day, sep="-"))]
```

```{r week}
met[, table(week(ymd))]
```
Keep observations of only first week of the month.
```{r first week only}
met_week1 <- met[week(ymd) == 31]
```

Compute variable means by station:
```{r means}
met_avg <- met_week1[, .(
  temp      = max(temp,na.rm=T),
  rh        = max(rh,na.rm=T), 
  wind.sp   = max(wind.sp,na.rm=T), 
  vis.dist  = max(vis.dist,na.rm=T), 
  dew.point = max(dew.point,na.rm=T), 
  lat       = max(lat,na.rm=T), 
  lon       = max(lon,na.rm=T), 
  elev      = max(elev,na.rm=T)
), by = "USAFID"]
```
```{r}
met_avg
```

```{r}
met_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                    fifelse(lon < -98 & lat > 39.71, "NW",
                    fifelse(lon < -98 & lat <= 39.71, "SW", "SE")))]
table(met_avg$region)
```

```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```


# 3. Use geom_violin to examine the wind speed and dew point temp by region
```{r dew violin plot}
met_avg[!is.na(region)] %>%
  ggplot() +
  geom_violin(mapping = aes(y = dew.point, x = 1, color=region, fill = region)) +
  facet_wrap(~ region, nrow=1)
```
```{r windspeed violin plot}
met_avg[!is.na(region) & !is.na(wind.sp)] %>%
  ggplot() +
  geom_violin(mapping = aes(x=1, y=wind.sp, color=region, fill=region)) +
  facet_wrap(~ region, nrow = 2)
  
```
# 4. Use geom_point with geom_smooth to examine the association between dew point temp and wind speed

Color points by region and make sure to deal w NA category.  Fit a linear regression line by region and describe graph observations.
```{r dew windsp point graph}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y = dew.point, color=region)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2)
```
# 5. Create barplots of weather stations by elevation category colored by region
```{r elev bar}
met_avg[!is.na(elev)] %>%
  ggplot()+
  geom_bar(mapping = aes(x = elev_cat, fill=region))
```


# 6. Use stat_summary to examine mean dew point and wind speed by region with std dev error bars
```{r dew windsp errorbar}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) + 
  stat_summary(fun.data=mean_sdl, geom = 'errorbar') +
  stat_summary(fun.data = mean_sdl, geom = "pointrange")
```
# 7. Generate a map of weather stations and show trend in relative humidity of top 10

```{r color palette}
library(leaflet)
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
```
```{r}
top10rh <- met_avg[ rank(-rh) <= 10]
```

```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid.', opacity=1)
rhmap
```

